---
title: "Boatable Days & Drought"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    logo: www/lynker_logo_white_transparent.png
    css: www/custom.css
runtime: shiny
---
<style>

.navbar, [data-toggle=tab], .navbar-brand  {   background-color:#1c3a5d;   border-color:black;   color:white; }

</style>

```{r}
# old YAML header and value box CSS 
# ---
# title: "Boatable Days & Drought"
# output: 
#   flexdashboard::flex_dashboard:
#     theme:
#       version: 4
#       bootswatch: yeti
#     orientation: columns
#     vertical_layout: fill
# runtime: shiny
# ---
# <style type="text/css">
# .value-box .value {
#     color: white;
# }
# .value-box .caption {
#     color: white;
# }
# </style>
```

```{r setup, include=FALSE}
library(flexdashboard)

# Shiny & Flexdashboard packages
library(shiny)
library(flexdashboard)
library(plotly)
library(shinyBS)
library(ggtext)
library(DT)

# Data manipulation
library(tidyr)
library(dplyr)
# library(tidyverse)
library(ggplot2)
# library(DT)
library(sf)
library(ggpmisc)
library(stringr)
library(ggrepel)
# library(grob)
# library(ggpubr)

# Mapping packages
library(leaflet)
library(leafem)

# water data
# library(nhdplusTools)
# library(cdssr)

# Load function data_utils.R file
source('R/utils.R')
# source('R/get_lm_relationships.R')
```


```{r, include = FALSE}
# path name
data_path <- "data/dashboard_data"

# read in sf site locations
gauge_sf <- st_read(paste0(data_path, "/tidy/gauge_sf.shp"))

# read in climate/drought data

bdays_climate <- readRDS(paste0(data_path, "/bdays_climate_data.rds"))

# unique site names
sites <- unique(bdays_climate$name_description)

# read in flow data
flow_thresholds <- read.csv(paste0(data_path, "/tidy/flow_thresholds.csv"))

# read in linreg results
linreg <- readRDS(paste0(data_path, "/tidy/linreg.rds"))

# slider range table
table <- data.frame(
  var = c("spei1y_may", "spi1y_may", "pdsi_may"),
  min = c(-2.25, -2.25, -5),
  max = c(2.25, 2.25, 5)
  )

```

```{r, include = FALSE}

# render leaflet map on page 1 when the app starts
output$map       <- leaflet::renderLeaflet({ basemap(shp = gauge_sf) })

# render the leaflet map on page 2 when the app starts
output$map2       <- leaflet::renderLeaflet({ basemap(shp = gauge_sf) })


```

Historical Data
======================================

Inputs {.sidebar}
-------------------------------------

**Historical Data**

This tab shows  historical streamflow and total number of summer season boatable days for 35 USGS Sites across Colorado. Boatable days are the number of days that the flow is between a specific flow range. Each USGS gage corresponds to a nearby American Whitewater stream segment with an established runnable range. 

Click on a USGS gage on the map to view.

**Gage Name:**
```{r gage_name1}
getname <- reactive({
  sprintf(usgs_name())
})

renderText(
  if(is.null(usgs_name())) {
    
    return(NULL)
    
  } else {
    {getname()}
})

```
\n

**River Segment Name:**
```{r segname1}
get_seg <- reactive({
  sprintf(usgs_name())
})

renderText(
  if(is.null(usgs_name())) {
    
    return(NULL)
    
  } else {
    {get_seg <- linreg %>%
    filter(site_name == usgs_name()) %>%
    slice(which.max(r2))
  seg <- get_seg$segment}
})
```


Column {data-width=500}
-----------------------------------------------------------------------
### Map
```{r}
leafletOutput('map')
```

```{r}
# old method for initially rendering leaflet 'map'
# # create map -------------
# output$map <- renderLeaflet({
# 
# 
#   leaflet(gauge_sf) %>%
#   addTiles() %>%
#   addCircleMarkers(label = gauge_sf$nm_dscr,
#                    layerId = ~nm_dscr)
# 
# })

# create reactive click variable -------------
# name of the USGS gauge based on the marker clicked on leaflet 'map'
usgs_name <- shiny::reactive({
   input$map_marker_click$id
})

# reactive value for highlighting a clicked marker from leaflet 'map'
selected_gauge <- shiny::reactive({
    # subset to point that matches the usgs_name2() that has been clicked on
    if(is.null(usgs_name())) {
      
      return(NULL)
      
    } else {
         gauge_sf[gauge_sf$nm_dscr == usgs_name(), ]
    }
  
})
```


Column {data-width=450}
-----------------------------------------------------------------------
### Streamflow Timeseries

```{r}
plotlyOutput("tsplot")

shinyBS::bsPopover(
  "tsplot",
  title     = "Streamflow Timeseries",
  content   = "This plot shows historical streamflow. The black horizontal lines are the upper and lower bounds of the runnable range that determines boatability",
  placement = "left",
  options   = list(container = "body")
  )
tags$style(HTML("
                .tooltip > .tooltip-inner {
                width: 400px;
                color: black;
                background-color: white;
                }
                "))
```

```{r context = "server"}
shiny::observeEvent(input$map_marker_click, {
  
  if(is.null(usgs_name())) {
    
    return(NULL)
    
  } else {
    
    print(paste0("usgs_name(): ", usgs_name()))
    
    print("Updating leaflet map w/ selected gauge...")
    
    # highlight the clicked on gauge in a different color
    leaflet::leafletProxy("map") %>%
      leaflet::clearMarkers() %>%
      leaflet::clearShapes() %>%
      leaflet::clearGroup(c("selected_gauge_group")) %>%
      leaflet::addCircleMarkers(
        data         = gauge_sf,
        color        = "black",
        opacity      = 0.7,
        fillColor    = "dodgerblue",
        fillOpacity  = 0.7,
        weight       = 2,
        stroke       = TRUE,
        label        = gauge_sf$nm_dscr,
        layerId      = ~nm_dscr
      ) %>%
      leaflet::addCircleMarkers(
        data         = selected_gauge(),
        color        = "black",
        opacity      = 1,
        fillColor    = "darkred",
        fillOpacity  = 1,
        weight       = 2,
        radius       = 12,
        stroke       = TRUE,
        group        = "selected_gauge_group",
        label        = selected_gauge()$nm_dscr,
        layerId      = ~nm_dscr
      )
    
    # output new timeseries plot ("tsplot") based on marker click on leaflet 'map'
    output$tsplot <-
      renderPlotly({
        df <- flow_thresholds %>%
          mutate(date = as.Date(date)) %>%
          filter(date >= "1980-01-01") %>%
          filter(station_nm == usgs_name())
        
        # grob1 <- grobTree(textGrob("Upper Bound", x=0.1,  y=0.95, 
        #                           hjust=0,
        #                           gp=gpar(col="darkred")))
        
        lastDate<- max(df$date)-1100
        Upper <- df$range_upper_bound[df$date==lastDate]
        Lower <- df$range_lower_bound[df$date==lastDate]

        
        ggplotly(
          ggplot(df, ) +
            geom_line(aes(x = date, y = flow), 
                      lwd = 0.8, color = "#4f77e3") +
            theme_bw() +
            geom_hline(yintercept = df$range_upper_bound[1], lwd = 0.5, color = "darkred") +
            geom_hline(yintercept = df$range_lower_bound[1], lwd = 0.5, color = "darkred") +
            # geom_text_repel(data=data.frame(),
            #           aes(x=c(lastDate, lastDate),
            #               y= c(Upper, Lower),
            #               label= c("Upper Bound", "Lower Bound"))) +
            # annotate(geom = 'text', label = 'sometext', 
            #          x = -Inf, y = Inf, hjust = 0, vjust = 1) +
            # geom_text(data=data.frame(), 
            #           aes(label = 'sometext', x = -Inf, y = Inf),
            #           hjust = 0, vjust = 1) +
            annotate(geom='text',
                     x=c(lastDate, lastDate),
                     y= c(Upper, Lower),
                     label= c("Upper Bound", "Lower Bound")) +
            # annotate("text", x = as.Date(-Inf, origin = '1970-01-01'), 
            #          y = Inf, hjust = 0, vjust = 0, label = "Hello world!") +
            labs(
              y = "Flow",
              x = "Date",
              title = usgs_name())
        )
      })
  }})

```



### Boatable Days
```{r}
plotlyOutput("bdaysplot")

shinyBS::bsPopover(
  "bdaysplot",
  title     = "Boatable Days",
  content   = "This barplot shows the total number of boatable days in a given year, based on a specific flow range",
  placement = "left",
  options   = list(container = "body")
  )
tags$style(HTML("
                .tooltip > .tooltip-inner {
                width: 400px;
                color: black;
                background-color: white;
                }
                "))

```

```{r context = "server"}

shiny::observeEvent(input$map_marker_click, {
  
  if(is.null(usgs_name())) {
    
    return(NULL)
    
  } else {
    
    output$bdaysplot <- renderPlotly({
      df <- bdays_climate %>%
        filter(name_description == usgs_name())
      ggplotly(
      ggplot(df,) +
        geom_bar(aes(x = year, y = boatable_days), stat = "identity") +
        theme_bw() +
        labs(
          y = "Boatable Days",
          x = "Date", 
          title = usgs_name()
        )
      )
    })
  }})
```


Prediction Tool
========================================


Inputs {.sidebar}
-------------------------------------
**Prediction Tool**

This tab shows the relationship between boatable days and common drought indices PDSI, SPEI, and SPI. The drought index that is shown for each gage is the one with the highest correlation with boatable days at that gage. Users can drag the slider below to show a predicted number of annual boatable days given a specific drought index value.

Click on a USGS gage on the map to view.

```{r context="server"}
# reactive value for highlighting a clicked marker
selected_gauge2 <- shiny::reactive({
    # subset to point that matches the usgs_name2() that has been clicked on
    if(is.null(usgs_name2())) {
      
      return(NULL)
      
    } else {
         gauge_sf[gauge_sf$nm_dscr == usgs_name2(), ]
    }
  
})
```


**Gage Name:**
```{r gage_name2}
getname2 <- reactive({
  sprintf(usgs_name2())
})

renderText(
  if(is.null(usgs_name2())) {
    
    return(NULL)
    
  } else {
    {getname2()}
})

```
\n

**River Segment Name:**
```{r segname}
get_seg <- reactive({
  sprintf(usgs_name2())
})

renderText(
  if(is.null(usgs_name2())) {
    
    return(NULL)
    
  } else {
    {get_seg <- linreg %>%
    filter(site_name == usgs_name2()) %>%
    slice(which.max(r2))
  seg <- get_seg$segment}
})
```


**Drought Index:**
```{r printvar}
get_var <- reactive({
  sprintf(best_var())
})

renderText(
  if(is.null(usgs_name2())) {
    
    return(NULL)
    
  } else {
    {get_var <- case_when(
                best_var() == "spi1y_may" ~ "May SPI 1 year",
                best_var() == "spei1y_may" ~ "May SPEI 1 year",
                TRUE ~ "May PDSI")}
    
    
})
```

<br> 

*Move slider to change Drought Index value*
```{r}
# initial slide input
sliderInput(
    "obs",
    label = NULL,
    value = 0,
    min = -5,
    max = 5,
    step = 0.25
    # step = NULL
 )

# set up observer to watch for the event when "best_var()" changes
# If best_var() changes (i.e. there is a new usgs_name2() value because one of the map markers was clicked),
#       then the sliderInput() is updated using the slider_min() and slider_max() reactive values that change based on the current "best_var()" 
shiny::observeEvent(best_var(), {
  shiny::updateSliderInput(
                session, 
                inputId = "obs",
                label   = NULL,
                value   = 0,
                min     = slider_min(),
                max     = slider_max(),
                step    = 0.25
                # step    = slider_step()
                # step = NULL
                )
  })


```

Column {data-width = 500}
-----------------------------------------------------------------------
### Map
```{r}
leafletOutput('map2')
```

```{r}
# old method for initially rendering leaflet 'map2'
# # create map -------------
# output$map2 <- renderLeaflet({
# 
# 
#   leaflet(gauge_sf) %>%
#   addTiles() %>%
#   addCircleMarkers(label = gauge_sf$nm_dscr,
#                    layerId = ~nm_dscr)
# 
# 
# })

# create reactive click variable -------------
# reactive value that changes based on the marker clicked on 'map2'
usgs_name2 <- shiny::reactive({
   input$map2_marker_click$id
})

# reactive value for highlighting a clicked marker from 'map2'
selected_gauge2 <- shiny::reactive({
    # subset to point that matches the usgs_name2() that has been clicked on
    if(is.null(usgs_name2())) {
      
      return(NULL)
      
    } else {
         gauge_sf[gauge_sf$nm_dscr == usgs_name2(), ]
    }
  
})
```



Column {data-width=450}
-----------------------------------------------------------------------


```{r r2value}
# get_r2 <- reactive({
#   sprintf(best_var())
# })
# 
# renderText(
#   if(is.null(usgs_name2())) {
#     
#     return(NULL)
#     
#   } else {
#     {forecast_data <- linreg %>%
#          filter(
#            site_name == usgs_name2(),
#            climate_var == best_var())
#     r2 <- paste0("R² = ", round(forecast_data$r2, 2))}
# })
```

```{r equation}
# old 'best_var()' reactive value --> this should work fine I just changed it slightly while troubleshooting
# # choose best variable to represent each site
# best_var <- shiny::reactive({
#   best_setup <- linreg %>%
#     filter(site_name == usgs_name2()) %>%
#     slice(which.max(r2))
#   print(paste0("best: ", best_setup$climate_var))
#   best <- best_setup$climate_var
# })

# choose best variable to represent each site
best_var <- shiny::reactive({
  if(is.null(usgs_name2())) {
    print(paste0('usgs_name2() is NULL'))
    return(NULL)
  }
  best_setup <- linreg %>%
    filter(site_name == usgs_name2()) %>%
    slice(which.max(r2))
  best <- best_setup$climate_var[1]
  
  print(paste0("best: ", best))
  
  return(best)
})

# slider input MINIMUM value based on "best_var()" reactive value
slider_min <- shiny::reactive({
    min_val <- table$min[table$var == best_var()]
    print(paste0("min_val: ", min_val))
    return(min_val)
})

# slider input MAXIMUM value based on "best_var()" reactive value
slider_max <- shiny::reactive({
    max_val <- table$max[table$var == best_var()]
    print(paste0("max_val: ", max_val))
    return(max_val)
})

# slider input step value based on "best_var()" reactive value,
#   if best_var() is pdsi_may, then the step = 1, otherwise step = 0.5
# (LOOKS LIKE THIS WON'T BE USED BUT COULD POTENTIALLY BE provided to the updateSliderInput() function "step=" argument
slider_step <- shiny::reactive({

    # set step = 1 if best_var() is "pdsi_may", otherwise set step = 0.5
    if(best_var() == "pdsi_may") {
      
      step <- 1
    
    } else {
      
      step <- 0.5
      
    }
    # print(paste0("step: ", step))
    return(step)
    
})

# prediction
y <- shiny::reactive({
  forecast_data <- linreg %>%
         filter(
           site_name == usgs_name2(),
           climate_var == best_var())
  
  slope <- forecast_data$m

  intercept <- forecast_data$intercept

  bdays_forecast <- slope*input$obs + intercept
}) 
```


### Scatterplot
```{r}
plotlyOutput("scatterplot")

shinyBS::bsPopover(
  "scatterplot",
  title     = "Linear Regression",
  content   = "This scatter plot shows the historical relationship between boatable days and the drought index with the highest predictive power. The red dot corresponds to the drought index value from the slider in the left-hand panel",
  placement = "left",
  options   = list(container = "body")
  )
tags$style(HTML("
                .tooltip > .tooltip-inner {
                width: 400px;
                color: black;
                background-color: white;
                }
                "))
```

```{r context = "server"}
# watch for marker clicks on map2
shiny::observeEvent(input$map2_marker_click, {
  # if usgs_name2() is NULL, then don't do anything
  if(is.null(usgs_name2())) {
    
    return(NULL)
  
  # otherwise if a valud usgs_name2(), then highlight the clicked gauge and update the plotly scatterplot
  } else {
    
    # print statements I was using for trouble shootingS
    print(paste0("usgs_name2(): ", usgs_name2()))
    print(paste0("best_var(): ", best_var()))
    print(paste0("slider_min(): ", slider_min()))
    print(paste0("slider_max(): ", slider_max()))
    print(paste0("slider_step(): ", slider_step()))
    
    print("Updating leaflet map w/ selected gauge...")

          # highlight the clicked on gauge in a different color
          leaflet::leafletProxy("map2") %>%
              leaflet::clearMarkers() %>%
              leaflet::clearShapes() %>%
              leaflet::clearGroup(c("selected_gauge_group2")) %>%
              leaflet::addCircleMarkers(
                data         = gauge_sf,
                color        = "black",
                opacity      = 0.7,
                fillColor    = "dodgerblue",
                fillOpacity  = 0.7,
                weight       = 2,
                stroke       = TRUE,
                label        = gauge_sf$nm_dscr,
                layerId      = ~nm_dscr
                ) %>%
              leaflet::addCircleMarkers(
                data         = selected_gauge2(),
                color        = "black",
                opacity      = 1,
                fillColor    = "darkred",
                fillOpacity  = 1,
                weight       = 2,
                radius       = 12,
                stroke       = TRUE,
                group        = "selected_gauge_group2",
                label        = selected_gauge2()$nm_dscr,
                layerId      = ~nm_dscr
                )
          
    # output plotly plots based on map2 marker click (usgs_name2() reactive value)
    output$scatterplot <-
      renderPlotly({
        df <- bdays_climate %>%
          filter(name_description == usgs_name2()) %>%
          select(name_description, id, boatable_days, best_var(), year)
        
        
        ggplotly(
          ggplot(df, 
                 aes(x = .data[[best_var()]], 
                     y = boatable_days, label = year)) +
            stat_poly_line() +
            # geom_smooth(method = "lm", formula = formula) +
            geom_point(size = 1.6)  +
            # stat_smooth(method = "lm", col = "black",
            # # se = FALSE,
            # lty = 2, lwd = 0.5) +
            theme_bw() +
            theme(plot.title = element_text(size = 10)) +
            labs(
              title = str_wrap(usgs_name2(), 35),
              y = "Boatable Days",
              x = case_when(
                best_var() == "spi1y_may" ~ "May SPI 1 year",
                best_var() == "spei1y_may" ~ "May SPEI 1 year",
                TRUE ~ "May PDSI"
              )
            ) +
            geom_point(aes(x = input$obs, y = y(), label = "Prediction"),
                       size = 3,
                       color = "darkred"
            ),
          tooltip = c("label")
        )
      })
  }})
```


### Number of Boatable Days

```{r}
# output the boatable days value box
valueBoxOutput("bdays_box")

shinyBS::bsPopover(
  "bdays_box",
  title     = "Number of Boatable Days",
  content   = "This is the number of boatable days predicted by the linear regression equation (red dot in plot above), given the drought index value indicated by the position of the slider tool in the left panel.",
  placement = "top",
  options   = list(container = "body")
  )
tags$style(HTML("
                .tooltip > .tooltip-inner {
                width: 400px;
                color: black;
                background-color: white;
                }
                "))
```

```{r}



# boatable dates value box rendering and calculations
output$bdays_box <- renderValueBox({
  
  # if usgs_name2() is NULL, render valueBox with nothing (NULL)
  if(is.null(usgs_name2())) {
  
      valueBox(
  
      paste(NULL)
  
      )
    
    # otherwise if usgs_name2() is NOT NULL (i.e. a click has happened)
    } else {
      
    # filter down 'linreg' model result dataframe to the specific site_name and the best_var()
    forecast_data <- linreg %>%
             filter(
               site_name == usgs_name2(),
               climate_var == best_var())
    
    # filter the boatable days climate data down to site_name == "usgs_name2()" 
    # then pivot climate variables from "pr" to the end of the dataframe
    # then do a filter again for just the "best_var()" values in the 'climate_var' column
    bdays_data <- 
      bdays_climate %>%
      filter(
        name_description == usgs_name2()
        ) %>%
      pivot_longer(
        cols      = pr:ncol(bdays_climate),
        names_to  = "climate_var",
        values_to = "value"
        ) %>%
      filter(climate_var == best_var())
    
    # count the number of boatable days (sum the values that are NOT NA)
    num_b_years <- sum(!is.na(bdays_data$boatable_days))
    
    # extract the slope
    slope <- forecast_data$m
    
    # extract the intercept
    intercept <- forecast_data$intercept
    
    # boatable days forecast equation
    bdays_forecast <- slope*input$obs + intercept
    
    # calculate standard error
    se <- forecast_data$sd/
      sqrt(num_b_years)
  
  # r2 <- forecast_data$r2
    
    # output value box with the number of boatable days
    valueBox(
  
      paste(round(bdays_forecast,0)),
      color = "#1c3a5d"
  
      )}
    })

```

### R²

```{r}
# output the boatable days value box
valueBoxOutput("bdays_box2")

shinyBS::bsPopover(
  "bdays_box2",
  title     = "R²",
  content   = "The R² value of the model for a given site and drought index. The closer the R² value is to 1.0, the higher the predictive power of the model.",
  placement = "top",
  options   = list(container = "body")
  )
tags$style(HTML("
                .tooltip > .tooltip-inner {
                width: 400px;
                color: black;
                background-color: white;
                }
                "))
```

```{r}

# boatable dates value box rendering and calculations
output$bdays_box2 <- renderValueBox({
  
  # if usgs_name2() is NULL, render valueBox with nothing (NULL)
  if(is.null(usgs_name2())) {
  
      valueBox(
  
      paste(NULL)
  
      )
    
    # otherwise if usgs_name2() is NOT NULL (i.e. a click has happened)
    } else {
      
    # filter down 'linreg' model result dataframe to the specific site_name and the best_var()
    forecast_data <- linreg %>%
             filter(
               site_name == usgs_name2(),
               climate_var == best_var())
  
    r2 <- forecast_data$r2
    
    # output value box with the number of boatable days
    valueBox(
  
      paste(round(r2,2)),
      color = "#1c3a5d"
  
      )}
    })

```

Table
=============================
```{r}
# DT::DTOutput("table")
```


```{r}
linreg <- linreg[with(linreg, order(river, -r2)),]

rownames(linreg) <- seq(1,length(linreg$river),1)


DT::datatable(linreg, options = list(
  bPaginate = FALSE
)) %>%
  DT::formatRound(columns = c("r2", "variance", "sd", "intercept", "m", "se"),  digits = 2)

```

